Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-10-13 manu-draft.qmd
update figure 2 for rn comments

A synthetic oxygen time series was then generated using the inputs from Apalachicola Bay ([@fig-synapa]) and @eq-ebase as a forward calculating model with estimated values for $a$, $R$, and $b$. All metabolic rates for production, respiration, and gas exchange ([@fig-synapa]b and c) were also estimated with the synthetic time series. The time series was estimated using the observed data described above, the separate metabolic components and their parameters in @eq-ebase, and approximate metabolic rates from the Odum open-water method as implemented in the WtRegDO R package [@Beck21] as an approximate estimate for back-calculation of the EBASE parameters. In short, $P$, $R$, and $D$ as $\text{mmol}/\text{m}^2/\text{d}$ were estimated using the Odum method and the separate synthetic components for EBASE were created by estimating the corresponding parameters, such that $P /PAR = a$, $R$ as is, and $D = k_w\left(C_{sat} - C \right)$, where $k_w$ includes the $b$ parameter. Daily estimates for each were created as the daily average for each parameter fit with a polynomial spline to create a smoothed output. The synthetic dissolved oxygen time series was then created as the integrated sum of the separate components for $P$, $R$, and $D$ divided by water column depth. The only parameter that could not be created as a continuous time series was $b$, which was set to 0.251 $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$ [@Wanninkhof14]. Accordingly, the synthetic dissolved oxygen time series and the observed water quality and meteorological data were used with EBASE to determine if reasonable estimates for the known parameters and metabolic rates were recovered.

@@

![Synthetic time series for one year of continuous data at Apalachicola Bay. Synthetic input to EBASE was the (a) dissolved oxygen time series, with results compared to the (b) synthetic $a$ parameter and (c) synthetic metabolic estimates for gross primary production ($P$), the negative of ecosystem respiration ($-R$), and gas exchange ($-D$).](figs/synapa.png){#fig-synapa}

>>>>

A synthetic oxygen time series was then generated using the inputs from Apalachicola Bay ([@fig-synapa]) and @eq-ebase as a forward calculating model with estimated values for $a$, $R$, and $b$. All metabolic rates for production, respiration, and gas exchange ([@fig-synapa]b) were also estimated with the synthetic time series. The time series was estimated using the observed data described above, the separate metabolic components and their parameters in @eq-ebase, and approximate metabolic rates from the Odum open-water method as implemented in the WtRegDO R package [@Beck21] as an approximate estimate for back-calculation of the EBASE parameters. In short, $P$, $R$, and $D$ as $\text{mmol}/\text{m}^2/\text{d}$ were estimated using the Odum method and the separate synthetic components for EBASE were created by estimating the corresponding parameters, such that $P /PAR = a$, $R$ as is, and $D = k_w\left(C_{sat} - C \right)$, where $k_w$ includes the $b$ parameter. Daily estimates for each were created as the daily average for each parameter fit with a polynomial spline to create a smoothed output. The synthetic dissolved oxygen time series was then created as the integrated sum of the separate components for $P$, $R$, and $D$ divided by water column depth. The only parameter that could not be created as a continuous time series was $b$, which was set to 0.251 $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$ [@Wanninkhof14]. Accordingly, the synthetic dissolved oxygen time series and the observed water quality and meteorological data were used with EBASE to determine if reasonable estimates for the known parameters and metabolic rates were recovered.

@@

![Synthetic time series for one year of continuous data at Apalachicola Bay. Synthetic data were created using the (a) $a$ parameter to estimate (b) metabolic estimates for gross primary production ($P$), the negative of ecosystem respiration ($-R$), and gas exchange ($-D$) and the (c) dissolved oxygen time series (including saturation, $C_{sat}$).](figs/synapa.png){#fig-synapa}
